Evaluating Staffing Characteristics, Healthcare-associated Infections, and Provider Characteristics in U.S. Nursing Homes, 2021

### 2023-05-05 code/analysis_code/analysis_paper.qmd
Project Analysis Thursday PM

f2

figure2 = here("results","Staff_prpd_distribution.png")

ggsave(filename = figure2, plot=f2)

@@

@@

figure_file = here("results","RiskStandardized_HAI_Rate_distribution.png")
ggsave(filename = figure1, plot=p1)

@@

###### Save for later ######
mytitlef4b <- "Figure: Relationship of adjusted HAI rate with Staffing Rating by nursing home"
vp4b <- ggplot(mydata, aes(x=as.factor(Staffing.Rating), y=HAI_039_RS_Rate)) + labs(x = " ", y = "Adjusted HAI Rate", title=  str_wrap(mytitlef4b, 75))+ geom_violin(trim=FALSE) +  stat_summary(fun = "mean",
               geom = "point",
               color = "red")
vp4b

@@

mydata$Staffing.Rating.F <- as.factor(mydata$Staffing.Rating)

@@

recipe_2 <- recipe(HAI_039_RS_Rate ~ Staffing.Rating  + Ownership.Type + Average.Number.of.Residents.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data= train_data)

@@

#Looking at model output on all of the data
wflow2 %>%  fit(data = mydata) %>% extract_fit_parsnip() %>%  tidy()
```

>>>>

@@

mydata$Staffing.Rating.F <- as.factor(mydata$Staffing.Rating)
mydata$owner <- substr(mydata$Ownership.Type, 1, 10)

@@

figure_file = here("results","RiskStandardized_HAI_Rate_distribution.png")
ggsave(filename = figure_file, plot=p1)

@@

vp4a

@@

@@

recipe_2 <- recipe(HAI_039_RS_Rate ~ Staffing.Rating  + owner + Average.Number.of.Residents.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data= train_data)

@@

mytitlescat <- "Observed and Predicted HAI Rates from Test Model"
scatter_model <- ggplot(model_aug2, aes(x=.pred, y=HAI_039_RS_Rate)) + labs(x = "Predicted Rate ", y = "Adjusted HAI Rate", title=  str_wrap(mytitlescat, 75))+ geom_point() +geom_abline() + stat_cor(method="pearson", label.x=6, label.y=20)
scatter_model

figure5 = here("results","scatter_model.png")
ggsave(filename = figure5, plot=scatter_model)

#Looking at model output on all of the data
table2save<- wflow2 %>%  fit(data = mydata) %>% extract_fit_parsnip() %>%  tidy()
table2 = here("results", "table2.RDS")
saveRDS(table2save, file=table2) 
```

### 2023-05-05 code/analysis_code/exploratory_analysis.qmd
Project Analysis Thursday PM

```{r}
p4 <- mydata %>% ggplot(aes(x=as.factor(Ownership.Type), y=HAI_039_RS_Rate)) + geom_boxplot() + geom_smooth(method='lm')
plot(p4)
figure_file = here("results","staff_hai_ownership.png")
ggsave(filename = figure_file, plot=p4) 
```

@@

#Applying the model to unseen test data

predict(model_fit2, test_data)

model_aug2\<- augment(model_fit1, test_data)

#pull RS HAI Rate and predictions

temps \<- model_aug2 %\>% pull(HAI_039_RS_Rate)

pred \<- model_aug2 %\>% pull(.pred)

#run rmse

rmse_vec(temps, pred)

@@

recipe_1 <- recipe(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data= train_data)

recipe_2 <- recipe(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score + as.factor(Staffing.Rating), data= train_data)

recipe_3 <- recipe(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score + as.factor(Staffing.Rating), data= train_data)

>>>>

mydata %>% count(Ownership.Type)

mydata$owner <- substr(mydata$Ownership.Type, 1, 10)

aggregate(mydata$Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day, list(mydata$owner), FUN=mean, na.rm=TRUE)
aggregate(mydata$HAI_039_RS_Rate, list(mydata$owner), FUN=mean)

p4 <- mydata %>% ggplot(aes(x=as.factor(owner), y=HAI_039_RS_Rate)) + geom_boxplot() + geom_smooth(method='lm')
plot(p4)
figure_file = here("results","hai_ownership.png")
ggsave(filename = figure_file, plot=p4)

view(mydata)
p5 <- mydata %>% ggplot(aes(x=as.factor(owner), y=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day)) + geom_boxplot() + geom_smooth(method='lm')
plot(p5)
figure_file = here("results","staff_ownership.png")
ggsave(filename = figure_file, plot=p5)

@@

@@

recipe_1 <- recipe(HAI_039_RS_Rate ~ owner + Average.Number.of.Residents.per.Day + Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data= train_data)

recipe_2 <- recipe(HAI_039_RS_Rate ~ owner + Average.Number.of.Residents.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score + Staffing.Rating, data= train_data)

recipe_3 <- recipe(HAI_039_RS_Rate ~ owner + Average.Number.of.Residents.per.Day + Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score + Staffing.Rating, data= train_data)

### 2023-05-05 products/manuscript/Manuscript.qmd
Project Analysis Thursday PM

@@

@@

From all data, we identified 10,502 nursing homes with complete data to include in our analysis. Complete description of the cohort included in the study in the Supplementary Files.

Distributions of all variables are included in the Appendix.

@@

We compared the three models using RMSE to determine which model best fit our data in the training dataset. The comparable RMSE rates are shown in Table 2.

***#Add R Code here to show Table 2 (RMSE for different model choices 1,2,3)***

We selected model 1, containing only the staffing rating as our best fitting model.When we applied model 1 to our test data we found that the predicted and observed values were fairly similar (Figure 5) and the RMSE was 1.79.

***#Add R Code here to show Figure 5***

@@

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable)
```

@@

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*

>>>>

## Hypothesis

The hypothesis of this study is that nursing homes with higher staffing rates and higher five-star staffing rate will be associated with lower healthcare associated infections requiring hospitalizations. We further hypothesize that models will show that actual staffing rates will be a better variable for predicting infection rates than five-star staffing ratings.

@@

## Exploratory analysis

First, we looked through the distribution of all the variables we may utilize in our study. Visually, we examined histograms, box plots, and other representations of the distribution. We then looked at how our continuous variables were distributed by the categorical staffing star metric, or by the HAI rate categorization.

In our exploratory analysis, we noticed missing values for some variables. Otherwise values were plausible and in range.

@@

Our exploratory analysis showed us fairly normal distributions of the risk standardized HAI requiring hospitalization rate and staffing variables. We noticed expected and appropriate correlation with the continuous rate variables and their categorical rankings. For example, as the number of staffing hours per resident per day increased, the staffing star rating also increased.

For interpretation, we reduced the \>10 ownership categories into Non-profit, For-profit, and Government. We saw potentially lower HAI rates in non-profit (7.00) compared to for-profit (7.9) nursing homes, with government-owned homes falling in the middle (7.4). Conversely, staffing rates were highest in non-profit nursing homes (4.2) compared to for-profit nursing homes (3.5), again with government in the middle (3.8).

From all data, we identified 10,502 nursing homes with complete data to include in our analysis. Some of these nursing homes were missing some staffing data.

@@

In our exploratory analysis, we compared the three models using RMSE to determine which model best fit our data in the training dataset. We selected model 1, containing only the staffing rating as our best fitting model.When we applied model 1 to our test data we found that the predicted and observed values were fairly similar (Figure 5) and the RMSE was 1.79.

```{r}
#| label: fig-result5
#| fig-cap: "Scatter plot of observed versus predicted values in test data"
#| echo: FALSE
knitr::include_graphics(here("results","scatter_model.png"))
```

###

The parameter estimates from the model applied to the full dataset are shown in Table 2.

```{r}
#| label: Table 2
#| fig-cap: "Table showing parameter estimates from selected model on full dataset"
#| echo: FALSE
resulttable2=readRDS("../../results/table2.RDS")
knitr::kable(resulttable2)
```

@@

@@

This analysis provides evidence that the risk standardized HAI metric may be able to elucidate important differences in quality of care in a nursing home setting. Further, it suggests that the five star staffing rating may be more informative than the staffing rate alone at predicting nursing homes with higher HAI rates. Nursing homes with inadequate staff should be provided additional resources to help ensure quality staffing is available to help keep residents safer from unnecessary harm.
